---
title: "Partitions"
author: "Brenda Betancourt, Giacomo Zanella, Rebecca C. Steorts"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Partitions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

We present a small example from Betancourt, Zanella, and Steorts (2020), "Random Partitions Models for Microclustering Tasks" \emph{Journal of the American Statistical Association, Theory and Methods}, Minor Revision. The partitions package performs entity resolution with categorical variables using partition-based Bayesian clustering models.

Our goals include:

- Creating a synthetic data set
- Illustrating how the user can perform entity resolution using the partitions package
- Illustrating how the user can calculate standard evaluation metrics when a unique identifer is known. 

## Loading all Packages Needed

We first load all packages needed for this example. 

```{r}
# load all packages need
# set seed for reproducibility 
library('partitions')
set.seed(123)
```

## Creating a Toy Data Set

We first create a synthetic data set by assuming a fixed partition. We assume that there are a maximum of four clusters. We assume that there are 50 records within each cluster. We assume that each record has 5 fields of categorical varaibles. We assume that the there are 10 possible categories per field. We assume that the distortion probability for each field is 0.01. The synthetic data set produces duplicate records using the `SimData()` function. Our syntehtic data set contains 500 records, with 200 unique records. 


```{r}
# true partition to generate simulated data
# 50 clusters of each size, max cluster size is 4
true_L <- c(50,50,50,50)
# number of fields
nfields <- 5
# number of cartegories per field
ncat <- rep(10,5)
# distortion probability for the fields
true_beta <- 0.01
# generate simulated data
x <- SimData(true_L, nfields, ncat, true_beta)
# dimension of data set
dim(x)
```

## Partition Priors

The package contains the implemetation of four random partition models that are used to perform entity resolution as a clustering task: 

- Two traditional random partition models: 
1.  Dirichlet process (DP) mixtures 
2.  Pitman–Yor process (PY) mixtures. 
-  Two random partition models that exhibit the microclustering property, which are 
Exchangeable Sequences of Clusters (ESC) models, which are referred to as (and are defined in our paper): 
3.  The ESCNB model 
4.  The ESCD model

## Posterior Samples

In order to obtain posterior samples of the cluster assignments and the model parameters the user needs to specificy the following

- the data,
- the random partition Prior ("DP", "PY", "ESCNB" or "ESCD"),
- the burn-in period, 
- and the number of iterations for the Gibbs sampler to run. 

Let's investigate this for the synthetic data set! 

```{r, eval=TRUE}
# example of drawing from the posterior with the ESCD prior 
PosESCD <- SampleCluster(data=x, Prior="ESCD", burn=5, nsamples=10)
```

The output is a **list** with two elements: 

- `Z`: A matrix of size nsamples x N containing the samples of the cluster assignments.
- `Params`: A matrix of size nsamples x # of model parameters containing the samples of the model parameters. The columns named beta_1 to beta_L  correspond to the distortion probabilities of the fields in the data.

# ATTN: alpha, r, and p have not been defined for the user. 

Observe that each row corresponds to an iteration of the Gibbs sampler. Observe that each column corresponds to a record. Observe that we have 500 records and 10 Gibbs iterations, as expected. We can inspect the first five row and first 10 columns of the posterior output. 

```{r, eval=TRUE}
dim(PosESCD$Z)
PosESCD$Z[1:5,1:10]
```

In addition, we can inspect the samples of the model parameters. 

```{r}
head(PosESCD$Params)
```

Samples for the DP, PY, and ESCNB models can be similarly obtained as:

```{r, eval=FALSE}
PosDP <- SampleCluster(x, "DP", 5, 10)
PosPY <- SampleCluster(x, "PY", 5, 10)
PosESCNB <- SampleCluster(x, "ESCNB", 5, 10)
```

## Evaluation Metrics

If the ground truth for the partition of the data is available, the average False Negative Rate (FNR) and False Discovery Rate (FDR) over the posterior samples can be computed (for any model) using the `mean_fnr` and `mean_fdr` functions:


```{r, eval=TRUE}
true_M <- length(true_L)
true_K <- sum(true_L)
# true cluster assignments
id <- rep(1:true_K, times=rep(1:true_M, times=true_L))
# average fnr
mean_fnr(PosESCD$Z,id)
# average fdr
mean_fdr(PosESCD$Z,id)
```

Of course, in practice, one would want to run the sampler much longer in practice to calculate the error rates above. 

